Analysis \& Visualization of Call Center Data using Python

### 2018-10-01 paper.tex
fixed some Typos

\begin{abstract}
\textit{The city of South Bend, Indiana operates a call center that serves as a primary point of contact for citizens.  The call center handles topics for nearly all of the city's departments.  An open data portal, maintained by the city, contains several years worth of information.  An analysis of this data was conducted using Python 3.6.  The data was 
analyzed for patterns by time of year, department, and topics with varying methods.  The cleaned, manipulated, and explored data was then developed into an 
interactive dashboard using the Bokeh library.  In doing so, an interactive HTML file will be distributed to the city, which can then be utilized, modified, and possibly connected 
directly to the data source.  Upon completetion, this analysis identified insights that could be actionable by the city.  July was found to be the most voluminous month, which was shown to be statistically signficant through the use of a two-sided t-test.  Additionally, the departments of Solid Waste and Water Works, along with many of their sub-topics, had some of the highest call volumes.  When calls were analyzed by duration, however, many lower volume topics were shown to take longer to resolve.}
\end{abstract}

@@

The purpose of this research is to uncover trends and insights that may be of use to call center management.  Within the scope of this analysis, call duration can be seen as the ``efficiency" measure within this call center, particularly on a topical basis.  This metric is identified by Baraka et al. as one of the primary measures to assess call center success.\cite{baraka}  The authors identify other parameters, but analyzing this operation's success level was not within the scope of this paper.
\par
Call center and queueing analysis has been the focus of academic research for years.  Brown et al. provide an excellent overview of the Poisson distribution modeling customer arrivals, as well as the accompanying assumptions.  While arrival distribution is not analyzed in this paper, the authors' assumptions are, for the most part, assumed to be true in this analysis.  These center around the general assumption that the customers and operators are statistically identical and that they all act independently.\cite{brown}  Zhang, Hong, and Zhang also descibe the arrival process as a Poisson distribution, but discuss models which may be more accurate alternatives.\cite{zhang}  They maintain some of the main assumptions as Brown.  As a result, differences in time-dependent parameters, customer attitudes and preferences, and operator skill levels are treated as neglegible.
\par
Call centers are a prime opportunity to utilize simulation techniques.  In fact, according to Bapat and Pruitte, simulation is the preferred method to analyze and determine the effectiveness of a call center.\cite{bapat}  It allows for evaluation of metrics beyond what a basic analysis encompasses.  For example, the scheduling of call agents should be optimized against call duration and abandonment, as documented by Saltzman and Mehrota.\cite{saltzman}  This could be viewed as a logical next step for this project, with this work discovering the baseline knowedge necessary to construct such a model.

@@

\textbf{Calls by topic --} were also analyzed by duration and volume.  When ranked by average duration, many obscure topics rose to the extremes.  Of the top 20 topics with the shortest duration, 19 had a total number of calls of less than three.  When viewed as a percentage of the total 202,500 call dataset, these calls are such a small proportion that it would give little actionable insight to the city.  A similar, though less extreme, situation occurred when viewing topics with longest duration.  In light of this exploration, it was determined that the most important results for the city to see would include only topics that composed a significant proportion of total calls.  When ranking by duration, only topics with total volume above the first quartile of total call volume were included.  This yielded results that could be viewed as more relevant to management.  Topics with long durations seemed to be centralized around complaints and billing issues, and quick topics were mostly comprised of general information inquiries.  When analyzed in this manner, there is a 195 second difference between the longest and shortest topics.
\par
Topic volume was dominated by topics within many of the top departments identified earlier in Figure 1.  The top two highest volume topics, by a large margin, were ``Miscellaneous Trash Information" and ``Water Miscellaneous".  This is reflective of positions of Solid Waste and Water Works in the departmental analysis.  In order to allow the city to effectively visualize this tandem effect, topics within the some of the busiest departments were plotted in an interactive jitterplot, as seen in Figure 4.

@@

While the above analysis yielded clear results, the call center did not have any specific questions for which it was seeking answers.  Instead, this project was geared towards discovering trends, patterns, and insights; these were then provided to the city so that they could formulate specific questions based on the story that was told by this data.  In order to provide this to city managment, a visualization dashboard was chosn as the most effective medium.  The Bokeh package was used to create interactive visualizations, and allowed the final product to be packaged into a HTML file and sent to the call center management.  In all, the dashboard includes visualizations of:

>>>>

\begin{abstract}
\textit{The city of South Bend, Indiana operates a call center that serves as a primary point of contact for citizens.  The call center handles topics for nearly all of the city's departments.  An open data portal, maintained by the city, contains several years worth of information.  An analysis of this data was conducted using Python 3.6.  The data was 
analyzed for patterns by time of year, department, and topics with varying methods.  The cleaned, manipulated, and explored data was then developed into an 
interactive dashboard using the Bokeh library.  In doing so, an interactive HTML file will be distributed to the city, which can then be utilized, modified, and possibly connected 
directly to the data source.  Upon completion, this analysis identified insights that could be actionable by the city.  July was found to be the most voluminous month, which was shown to be statistically significant through the use of a two-sided t-test.  Additionally, the departments of Solid Waste and Water Works, along with many of their sub-topics, had some of the highest call volumes.  When calls were analyzed by duration, however, many lower volume topics were shown to take longer to resolve.}
\end{abstract}

@@

The purpose of this research is to uncover trends and insights that may be of use to call center management.  Within the scope of this analysis, call duration can be seen as the ``efficiency" measure within this call center, particularly on a topical basis.  This metric is identified by Baraka et al. as one of the primary measures to assess call center success.\cite{baraka}  The authors identify other parameters, but analyzing this operation's success level was not within the scope of this paper.
\par
Call center and queuing analysis has been the focus of academic research for years.  Brown et al. provide an excellent overview of the Poisson distribution modeling customer arrivals, as well as the accompanying assumptions.  While arrival distribution is not analyzed in this paper, the authors' assumptions are, for the most part, assumed to be true in this analysis.  These center around the general assumption that the customers and operators are statistically identical and that they all act independently.\cite{brown}  Zhang, Hong, and Zhang also describe the arrival process as a Poisson distribution, but discuss models which may be more accurate alternatives.\cite{zhang}  They maintain some of the main assumptions as Brown.  As a result, differences in time-dependent parameters, customer attitudes and preferences, and operator skill levels are treated as negligible.
\par
Call centers are a prime opportunity to utilize simulation techniques.  In fact, according to Bapat and Pruitte, simulation is the preferred method to analyze and determine the effectiveness of a call center.\cite{bapat}  It allows for evaluation of metrics beyond what a basic analysis encompasses.  For example, the scheduling of call agents should be optimized against call duration and abandonment, as documented by Saltzman and Mehrota.\cite{saltzman}  This could be viewed as a logical next step for this project, with this work discovering the baseline knowledge necessary to construct such a model.

@@

\textbf{Calls by topic --} were also analyzed by duration and volume.  When ranked by average duration, many obscure topics rose to the extremes.  Of the top 20 topics with the shortest duration, 19 consisted of less than three total calls.  When viewed as a percentage of the total 202,500 call dataset, these calls are such a small proportion that it would give little actionable insight to the city.  A similar, though less extreme, situation occurred when viewing topics with longest duration.  In light of this exploration, it was determined that the most important results for the city to see would include only topics that composed a significant proportion of total calls.  When ranking by duration, only topics with total volume above the first quartile of total call volume were included.  This yielded results that could be viewed as more relevant to management.  Topics with long durations seemed to be centralized around complaints and billing issues, and quick topics were mostly comprised of general information inquiries.  When analyzed in this manner, there is a 195 second difference between the longest and shortest topics.
\par
Topic volume was dominated by topics within many of the top departments identified earlier in Figure 1.  The top two highest volume topics, by a large margin, were ``Miscellaneous Trash Information" and ``Water Miscellaneous".  This is reflective of positions of Solid Waste and Water Works in the departmental analysis.  In order to allow the city to effectively visualize this tandem effect, topics within the some of the busiest departments were plotted in an interactive jitterplot, as seen in Figure 4.

@@

While the above analysis yielded clear results, the call center did not have any specific questions for which it was seeking answers.  Instead, this project was geared towards discovering trends, patterns, and insights; these were then provided to the city so that they could formulate specific questions based on the story that was told by this data.  In order to provide this to city management, a visualization dashboard was chosen as the most effective medium.  The Bokeh package was used to create interactive visualizations, and allowed the final product to be packaged into a HTML file and sent to the call center management.  In all, the dashboard includes visualizations of:
